\chapter{Constraining Mesh to a Geometric Domain}
\label{sec:geom}

Vertex positions may be constrained to a geometric domain by providing Mesquite
with an optional instance of the \texttt{Mesquite::MeshDomain} interface.  This
interface provides two fundamental capabilities: mesh-geometry classification,
and interrogation of local geometric properties.  The methods defined in the 
\texttt{Mesquite::MeshDomain} interface combine both queries into single
operation.  Queries are passed a mesh entity handle (see Section \ref{sec:MeshData}),
and are expected to interrogate the geometric domain that the specified mesh
entity is classified to.

If Mesquite is used to optimize the mesh of a B-Rep solid model (the data model used by all modern CAD systems), then the domain is composed of geometric vertices,
curves, surfaces, and volumes.  Curves are bounded by end vertices, surfaces are
bounded by loops (closed chains) of curves, and volumes are bounded by groups of
surfaces.  Mesquite expects each surface element (triangle, quadrilateral, etc.)
to be associated with a 2D domain (surface).  Vertices may be associated with
a geometric entity that either contains adjacent mesh elements or bounds the
geometric entity containing the adjacent elements. Mesquite does not use 
geometric volumes.  A query for the closest location on the domain for a vertex 
or element whose classification is a geometric volume should simply return the 
input position.  

It is possible to define an optimization problem such that mesh classification
data need not be provided in a \texttt{Mesquite::MeshDomain} implementation. 
This is done by optimizing the mesh associated with each simple geometric
component of the domain separately, with the boundary vertices flagged as fixed.
The following pseudo-code illustrates such an approach for a B-Rep type geometric
domain:
\begin{verbatim}
for each geometric vertex
 make associated vertex as fixed
end-for
for each curve
 do any application-specific optimization of curve node placement
 mark associated mesh vertices as fixed
end-for
for each surface
 define Mesquite::MeshDomain for surface geometry
 invoke Mesquite to optimize surface mesh
 mark all associated mesh vertices as fixed
end-for
for each volume
 invoke Mesquite to optimize volume mesh w/o Mesquite::MeshDomain
end-for
\end{verbatim}


\section{The ITAPS iGeom and iRel Interfaces} \label{sec:ITAPS}

Mesquite can access mesh domain data through the \emph{iGeom} and \emph{iRel} interface defined by the ITAPS Work Group.  These interfaces provide APIs for accessing B-Rep geometric data and associating mesh with geometry (classifiation), respectively.  Mesquite provides the \texttt{Mesquite::MsqIGeom} class (\texttt{MsqIGeom.hpp}) as an adaptor for interfacing with appications that present the \emph{iGeom} and \emph{iRel} interfaces.  The use of the \emph{iRel} interface is optional.  If all the mesh vertices are constrained to a single geometric surface, it is sufficient to provide only an iGeom instance to \texttt{Mesquite::MsqIGeom}.  If vertices are constrained to different geometric entities, then the \emph{iRel} interface must be provided to \texttt{Mesquite::MsqIGeom} so Mesquite can determine which iMesh entity a given vertex is constrained to lie in.


\section{Simple Geometric Domains} \label{sec:MsqGeom}

Mesquite provides several implementations of the
\texttt{Mesquite::MeshDomain} interface for simple geometric primitives.  These
include
\begin{itemize}
\item \texttt{PlanarDomain}: An unbounded planar domain.
\item \texttt{SphericalDomain}: A closed spherical domain.
\item \texttt{CylinderDomain}: An unbounded cylindrical domain.
\item \texttt{BoundedCylinderDomain}: A bounded cylindrical domain.
\end{itemize}
The \texttt{PlanarDomain} is often used to map $\mathbb{R}^{2}$ optimization
problems to $\mathbb{R}^{3}$.  The others are used primarily for testing
purposes.  The \texttt{BoundedCylinderDomain} provides some simplistic
mesh-geometry classification capabilities.  The others do not provide any
classification functionality.  
